Correspondence Analysis of Two Mode Networks

### 2025-01-29 main-R1.tex

\begin{figure}[ht!]
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
    \includegraphics[width=1.0\textwidth]{Plots/p-reflections.png}
    \caption{Southern Women data with rows and columns ordered as in the original DGG table.}
    \label{fig:sw}
\end{figure}
\subsection{Empirical Example} \label{subsec:ex}
\subsubsection{Souther Women Data}
The following analyses use the classic Southern Women data \citep{davis1941} (DGG) as the running example, a dataset that, according to \citet[p.]{freeman2003finding}, ``has become something of a touchstone for comparing analytic methods in social network analysis,'' having been analyzed ``again and again,'' and reappearing ``whenever any network analyst wants to explore the utility of some new tool for analyzing data.'' Freeman himself would go to review twenty-one such attempts, and the list has only gotten longer since his analytic ``meta-analysis'' \citep[e.g.,][]{doreian2004generalized, field2006identifying, roffilli2006identifying, wang2009exponential, borgatti2011analyzing, lerner2022dynamic, batagelj2022analysis,  lizardo2024two}. As we can see, novel insights can still be garnered from these old data.

One key source of confusion in the mini-literature that exists around this dataset concern the order of presentation of the rows and columns. The reason for this is while the row mode (people) is ``nominal'' in both the scaling and the literal senses (being composed of women's names), the column mode (events) is more ambiguous. It \textit{could} be treated as nominal, but the original DGG table includes event \textit{dates}, which means that it could also be treated as having a natural ordering, which could be reflected in the way we list the columns.  As \citet{freeman2003finding} notes, in the canonical original table both the row and column modes had been re-ordered in presentation to reveal what DGG thought was the group structure among the women; this means that the events were \textit{not} presented in chronological order in terms of column ordering. To add to the confusion, the original DGG table added arbitrary ordered numerical designators to both people and events, such that people came to refer to "W1" or "W11" and "E1" or "E13" based on the original DGG ordering, which may not have mattered for the women, but it mattered for the events because "E14" in the original DGG data is not the last chronological even among the fourteen columns, but in fact, as shown in~\ref{tab:southern}, it is the event that occurred on August 3rd, and there were three events that occurred after that one (E4, E8, and E13) in the DGG numbering.

To further add to the confusion, some early analyses of the data, including Breiger's \citeyearpar{breiger1974duality} classic article and \citet{doreian1979evolution} used a \textit{different} ordering of the women (taken from \citet{homans}) from that of the original DGG table reproduced in \citet{freeman2003finding}, but with the events ordered chronologically, which differs from the original DGG Table reproduced in \citet{freeman2003finding}. This means that the arbitrary ordinal labeling of the rows (e.g., "W1, "W2", "W3", etc.) and the columns (e.g., "E1, "E2", "E3", etc.) in these (and other) publications---see e.g., \citep[table 1]{doreian1979evolution}---does not correspond to the DGG ordering reproduced in \citet{freeman2003finding}. Different versions of the dataset reproduce different versions of the labeling. Of course, at the end of the day, the specific labeling of the nodes does not really matter, since we are usually interested in the structural properties of the rows and columns. The even ordering however, does matter for analyses that try to exploit the temporal properties of the data. Less substantively, but still importantly, the labeling of both the rows and columns does matter when trying to reconcile claims across publications from the narrative included in various papers, since "W1" or "E8" in one publication (e.g., \citet{freeman2003finding}) may not correspond to correspond to "W1" or "E8" in the other (e.g., \citet{doreian1979evolution}).

>>>>

\begin{figure}[ht!]
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
    \includegraphics[width=0.65\textwidth]{Plots/southern-women.jpg}
    \caption{Southern Women data with rows and columns ordered as in the original DGG table.}
    \label{fig:sw}
\end{figure}

\subsection{Empirical Example} \label{subsec:ex}
\subsubsection{Souther Women Data}
The following analyses use the classic Southern Women data \citep{davis1941} (DGG) as the running example, a dataset that, according to \citet[p.]{freeman2003finding}, ``has become something of a touchstone for comparing analytic methods in social network analysis,'' having been analyzed ``again and again,'' and reappearing ``whenever any network analyst wants to explore the utility of some new tool for analyzing data.'' Freeman himself would go to review twenty-one such attempts, and the list has only gotten longer since his analytic ``meta-analysis'' \citep[e.g.,][]{doreian2004generalized, field2006identifying, roffilli2006identifying, wang2009exponential, borgatti2011analyzing, lerner2022dynamic, batagelj2022analysis,  lizardo2024two}. As we can see, novel insights can still be garnered from these old data.

One key source of confusion in the mini-literature that exists around this dataset concerns the order of presentation of the rows and columns. The reason for this is while the row mode (people) is ``nominal'' in both the scaling and the literal senses (being composed of women's names), the column mode (events) is more ambiguous. It \textit{could} be treated as nominal; however, the original DGG table includes event \textit{dates}, which means that it could also be treated as having a natural ordering, which could be reflected in the way we list the columns.  As \citet{freeman2003finding} notes, in the canonical original table, both the row and column modes had been re-ordered in a presentation to reveal what DGG thought was the group structure among the women; this means that the events were \textit{not} presented in chronological order from left to right across the columns. To add to the confusion, the original DGG table added arbitrary ordered numerical designators to both people and events, such that people came to refer to "W1" or "W11" and "E1" or "E13" based on the original DGG ordering, which may not have mattered for the women, but it mattered for the events because "E14" in the original DGG data is not the last chronological even among the fourteen columns. In fact, as shown in~\ref{tab:southern}, it is the event that occurred on August 3rd, and three events occurred after that (E4, E8, and E13) in the DGG numbering.

To further add to the confusion, some early analyses of the data, including Breiger's \citeyearpar{breiger1974duality} classic article and \citet{doreian1979evolution} used a \textit{different} ordering of the women (taken from \citet{homans}) from that of the original DGG table reproduced in \citet{freeman2003finding}, but with the events ordered \textit{chronologically}, which differs from the original DGG Table reproduced in \citet{freeman2003finding}. This means that the arbitrary ordinal labeling of the rows (e.g., "W1, "W2", "W3", etc.) and the columns (e.g., "E1, "E2", "E3", etc.) in these (and other) publications---see e.g., \citep[table 1]{doreian1979evolution}---does not correspond to the DGG ordering reproduced in \citet{freeman2003finding} and more recent publications \citep[e.g.,][]{borgatti2014analyzing, batagelj2022analysis}; different versions of the dataset reproduce different versions of the labeling. Of course, the specific labeling of the nodes does not really matter since we are usually interested in the structural properties of the rows and columns. Event ordering, however, matters for analyses that try to exploit the temporal properties of the data. Less substantively, but still importantly, the labeling of both the rows and columns does matter when trying to reconcile claims across publications from the narrative included in various papers, since "W1" or "E8" in one publication (e.g., \citet{freeman2003finding}) may not correspond to "W1" or "E8" in the other (e.g., \citet{doreian1979evolution}).
